\documentclass[12pt, a4paper]{report}

% --- PACKAGES ---
\usepackage{iftex}
\ifPDFTeX
  \usepackage[utf8]{inputenc}
  \usepackage[T1]{fontenc}
  \renewcommand{\rmdefault}{ptm}
\else
  \usepackage{fontspec}
  \setmainfont[
    Path=/System/Library/Fonts/Supplemental/,
    UprightFont={Times New Roman.ttf},
    BoldFont={Times New Roman Bold.ttf},
    ItalicFont={Times New Roman Italic.ttf},
    BoldItalicFont={Times New Roman Bold Italic.ttf}
  ]{Times New Roman}
\fi
\usepackage[margin=25mm]{geometry} % 25 mm on all sides (Annex B)
\usepackage{setspace}       % For line spacing
\usepackage{titlesec}       % To customize chapter headings
\usepackage{graphicx}       % For images
\usepackage{caption}        % For caption formatting
\usepackage{indentfirst}    % Indent first paragraph of sections
\usepackage{enumitem}       % For list formatting
\usepackage{tocloft}        % To customize Table of Contents
\usepackage{fancyhdr}      % For custom headers/footers
\pagestyle{plain}          % Bottom-center page numbering (Zalacznik B)

\usepackage{array}
\usepackage{booktabs}
\usepackage{float}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[english]{babel}
\usepackage{subcaption}
\usepackage{url}           % Safe URL handling in apacite bibliography entries
\usepackage{microtype}     % Safer line breaking and character protrusion
\usepackage{apacite}
\usepackage{etoolbox}
\usepackage{multirow}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}

\makeatletter
\newcommand{\PatchAPACciteBreaks}{%
  \renewcommand{\BBAY}{,\nobreakspace}%
  \renewcommand{\BBYY}{,\nobreakspace}%
  \patchcmd{\@ifauthorsunequalc@de}
    {{\csname b@\@citeb\APAC@extra@b@citeb\endcsname}}
    {\mbox{\csname b@\@citeb\APAC@extra@b@citeb\endcsname}}
    {}{\PackageWarning{masters_thesis}{Could not patch apacite author citation breaks}}%
  \renewenvironment{APACrefauthors}{%
    \begingroup\APACrefauthstyle
  }{%
    \endgroup\nobreak
  }%
}
\AtBeginDocument{\PatchAPACciteBreaks}
\makeatother

% --- CONFIGURATION ---

% 1. Line Spacing: 1.5 for main text
\onehalfspacing
\setlength{\emergencystretch}{2em}
\setlength{\intextsep}{\baselineskip}
\setlength{\textfloatsep}{\baselineskip}
\setlength{\floatsep}{\baselineskip}

% 2. Paragraph Indentation: 1 cm
\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

% 3. Chapter and section heading formatting
\renewcommand{\contentsname}{TABLE OF CONTENTS}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\refname}{REFERENCES}
\addto\captionsenglish{%
  \renewcommand{\contentsname}{TABLE OF CONTENTS}%
  \renewcommand{\bibname}{REFERENCES}%
  \renewcommand{\refname}{REFERENCES}%
}
\titleformat{\chapter}[display]
{\normalfont\normalsize\bfseries\centering\setstretch{1}}
{CHAPTER \Roman{chapter}}
{12pt}
{\MakeUppercase}
\titleformat{name=\chapter,numberless}[block]
{\normalfont\normalsize\bfseries\centering\setstretch{1}}
{}
{0pt}
{\MakeUppercase}
\titlespacing*{\chapter}{0pt}{0pt}{1.5\baselineskip}

\titleformat{\section}
{\normalfont\normalsize\bfseries\setstretch{1}}
{\thesection.}
{1em}
{}
\titleformat{\subsection}
{\normalfont\normalsize\bfseries\setstretch{1}}
{\thesubsection.}
{1em}
{}
\titlespacing*{\section}{0pt}{1.5\baselineskip}{1.5\baselineskip}
\titlespacing*{\subsection}{0pt}{1.5\baselineskip}{1.5\baselineskip}
\renewcommand{\thechapter}{\Roman{chapter}}
\renewcommand{\thesection}{\arabic{chapter}.\arabic{section}}
\renewcommand{\thesubsection}{\arabic{chapter}.\arabic{section}.\arabic{subsection}}
\renewcommand{\cfttoctitlefont}{\normalfont\normalsize\bfseries}
\renewcommand{\cftaftertoctitle}{}
\setlength{\cftbeforetoctitleskip}{0pt}
\setlength{\cftaftertoctitleskip}{1.5\baselineskip}
\makeatletter
\renewcommand{\@cftmaketoctitle}{%
  \addpenalty\@secpenalty
  \vspace*{\cftbeforetoctitleskip}%
  \@cftpagestyle
  {\interlinepenalty\@M
  \noindent\makebox[\linewidth][c]{\cfttoctitlefont\contentsname\cftaftertoctitle}%
  \cftmarktoc
  \par\nobreak
  \vskip \cftaftertoctitleskip
  \@afterheading}%
}
\makeatother
\newcommand{\tocchapterprefix}{\renewcommand{\cftchappresnum}{CHAPTER~}}
\tocchapterprefix
\setlength{\cftchapnumwidth}{7em}
\renewcommand{\cftchapfont}{\normalfont}
\renewcommand{\cftchappagefont}{\normalfont}

% 4. Captions, counters, equations, lists, and tables
\makeatletter
\@removefromreset{figure}{chapter}
\@removefromreset{table}{chapter}
\@removefromreset{equation}{chapter}
\makeatother
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\renewcommand{\theequation}{\arabic{equation}}
\captionsetup{
  font={normalsize,stretch=1},
  labelfont=bf,
  labelsep=period,
  justification=justified,
  singlelinecheck=false,
  skip=\baselineskip
}
\captionsetup[figure]{position=top}
\captionsetup[table]{position=top}
\setlist[itemize]{label=--, leftmargin=1cm, itemsep=0pt, topsep=0pt, parsep=0pt}
\setlist[enumerate]{leftmargin=1cm, itemsep=0pt, topsep=0pt, parsep=0pt}
\setlength{\arrayrulewidth}{0.4pt}
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.2}
\renewcommand{\toprule}{\hline}
\renewcommand{\midrule}{\hline}
\renewcommand{\bottomrule}{\hline}

% 5. Custom commands for notes and sources under figures/tables
\newcommand{\floatnote}[1]{%
  \par\vspace{0.5\baselineskip}%
  {\singlespacing\noindent\footnotesize\raggedright\textit{Note.} #1\par}%
  \vspace{0.5\baselineskip}%
}
\newcommand{\floatsource}[1]{%
  \par\vspace{0.5\baselineskip}%
  {\singlespacing\noindent\footnotesize\raggedright\textit{Source:} #1\par}%
}
\newcommand{\floatnotewithsource}[2]{%
  \par\vspace{0.5\baselineskip}%
  {\singlespacing\noindent\footnotesize\raggedright\textit{Note.} #1\\[-0.2\baselineskip]\textit{Source:} #2\par}%
}
\newcommand{\floatnotewithprob}[2]{%
  \par\vspace{0.5\baselineskip}%
  {\singlespacing\noindent\footnotesize\raggedright\textit{Note.} #1\\[-0.2\baselineskip]#2\par}%
}
\newcommand{\datasetsource}{\floatsource{Own elaboration based on the thesis dataset.}}
\newcommand{\ownsource}{\floatsource{Own elaboration.}}
\newcommand{\datasetnote}[1]{\floatnotewithsource{#1}{Own elaboration based on the thesis dataset.}}
\newcommand{\ownnote}[1]{\floatnotewithsource{#1}{Own elaboration.}}
\newcommand{\source}[1]{\floatsource{#1}}

\makeatletter
\newcommand{\compactlistheading}[1]{%
  \par\noindent{\normalfont\normalsize\bfseries #1}\par\nobreak
  \vspace{0.25\baselineskip}%
}
\newcommand{\combinedlistoftablesandfigures}{%
  \chapter*{LIST OF TABLES AND FIGURES}%
  \addcontentsline{toc}{chapter}{LIST OF TABLES AND FIGURES}%
  \begingroup
  \singlespacing
  \let\addvspace\@gobble
  \compactlistheading{Tables}%
  \@starttoc{lot}%
  \par\vspace{0.5\baselineskip}%
  \compactlistheading{Figures}%
  \@starttoc{lof}%
  \endgroup
}
\makeatother

% --- SECTIONING STYLE (Kept from original but adapted) ---
% \renewcommand*\thesection{\thechapter.\arabic{section}.}
% \renewcommand*\thesubsection{\thesection\arabic{subsection}.}

% --- METADATA ---
\newcommand{\thesisTitle}{Causal AI for Detecting Crypto Market Manipulation}
\newcommand{\thesisAuthor}{Igor Kołodziej}
\newcommand{\studentID}{440239}
\newcommand{\supervisor}{Dr Hab. Stanisław Cichocki, Prof. UW}
\newcommand{\fieldOfStudy}{Data Science and Business Analytics}
\newcommand{\submissionPlace}{Warsaw}
\newcommand{\submissionMonth}{May}
\newcommand{\submissionYear}{2026}

\begin{document}

% ===================================================================
% 1. TITLE PAGE
% ===================================================================
\begin{titlepage}
  \thispagestyle{empty}
  \setcounter{page}{0}
  \centering
  {\fontsize{18pt}{21.6pt}\selectfont University of Warsaw \par}
  {\fontsize{18pt}{21.6pt}\selectfont Faculty of Economic Sciences \par}

  \vspace{2.5cm}

  {\fontsize{14pt}{16.8pt}\selectfont \thesisAuthor \par}
  {\fontsize{14pt}{16.8pt}\selectfont Student's book no.: \studentID \par}

  \vspace{1.8cm}

  {\bfseries\fontsize{20pt}{24pt}\selectfont \thesisTitle \par}

  \vspace{1.8cm}

  {\normalsize Master's thesis \par}
  {\normalsize on the field of: \fieldOfStudy \par}

  \vspace{1.8cm}

  \begin{flushright}
    The thesis written under supervision of\par
    \supervisor\par
    Faculty of Economic Sciences\par
  \end{flushright}

  \vfill

  {\normalsize \submissionPlace, \submissionMonth\ \submissionYear \par}
\end{titlepage}
\setcounter{page}{1}

% ===================================================================
% 2. STATEMENTS
% ===================================================================
\thispagestyle{plain}

\noindent \textbf{Statement of the Supervisor on Submission of the Thesis}

\vspace{0.45cm}

\noindent I hereby certify that the thesis submitted has been prepared under my supervision and I declare that it satisfies the requirements of submission in the proceedings for the award of a degree.

\vspace{1.25cm}

\noindent Date \hfill Signature of the Supervisor

\vspace{1.45cm}

\noindent \textbf{Statement of the Author on Submission of the Thesis}

\vspace{0.45cm}

\noindent Aware of legal liability I certify that the thesis submitted has been prepared by myself and does not include information gathered contrary to the law.

\vspace{0.35cm}

\noindent I also declare that the thesis submitted has not been the subject of proceedings resulting in the award of a university degree.

\vspace{0.35cm}

\noindent Furthermore I certify that the submitted version of the thesis is identical with its attached electronic version.

\vspace{1.15cm}

\noindent Date \hfill Signature of the Author of the thesis

\vspace{1.2cm}

\noindent \textbf{Declaration on the Use of Artificial Intelligence (AI) Systems in the Thesis}

\vspace{0.45cm}

\noindent During the preparation of this thesis, the author used ChatGPT, Grammarly, and Gemini to support proofreading and language editing, sentence rephrasing, grammar and style checking, assistance with LaTeX formatting, and coding-related consultation. Gemini was also used to support the annotation and contextual validation of Telegram pump-signal messages described in the empirical workflow. The author reviewed and verified the thesis content, code, and analytical outputs. The thesis represents the author's own intellectual contribution, and the author takes full responsibility for its final form.

\vspace{1.15cm}

\noindent Date \hfill Signature of the Author

\newpage

% ===================================================================
% 3. ABSTRACT & KEYWORDS
% ===================================================================
\thispagestyle{plain}

\begin{center}
  \textbf{Abstract}
\end{center}
\noindent Telegram-coordinated cryptocurrency pump-and-dump events are analyzed by linking signal features, high-frequency exchange data, and asset metadata. The empirical design estimates how \textit{Hype Intensity}, \textit{Urgency Index}, and \textit{Instruction Density} relate to the \textit{10-Minute Logarithmic~Return} using Double Machine Learning and heterogeneity checks. Estimated effects differ by market capitalization. \textit{Urgency Index} is most clearly positive among small-cap assets, while several effects weaken or turn negative in larger markets. An XGBoost model evaluates whether the binary \textit{Success}/\textit{Non-Success} outcome can be classified from information available before the reveal. The findings characterize pump signals as short-term coordination tools whose impact depends on market conditions.

\vspace{1cm}
\begin{center}
  \textbf{Keywords}
\end{center}
\begin{center}
  causal ai, cryptocurrency, market manipulation, pump-and-dump, double machine learning
\end{center}

\begin{center}
  (przyczynowa sztuczna inteligencja, kryptowaluty, manipulacja rynkowa, pump-and-dump, uczenie maszynowe)
\end{center}

\vspace{1cm}
\begin{center}
  \textbf{Tytuł pracy w języku polskim}
\end{center}
\begin{center}
  Przyczynowa SI w wykrywaniu manipulacji na rynkach kryptowalutowych
\end{center}

\newpage

% ===================================================================
% 4. TABLE OF CONTENTS
% ===================================================================
\tableofcontents
\newpage

\input{chapters/introduction}

\chapter{LITERATURE REVIEW}
\label{ch:literature-framework}

\input{chapters/chapter_1}

% --- PART 1: DATA DESCRIPTION ---
\graphicspath{{figures/}}

\input{chapters/chapter_2}


% --- PART 2: METHODOLOGY ---
\chapter{METHODOLOGY}

\input{chapters/chapter_3/chapter_3_methodology}

% --- PART 3: EMPIRICAL RESULTS ---
\chapter{EXPERIMENTAL RESULTS}
\label{ch:results}
\input{chapters/chapter_4/chapter_4_results}

% --- SECTION 4.3: DISCUSSION ---
\section{Discussion}
\label{sec:discussion}
\input{chapters/chapter_5/chapter_5_interpretation}

% --- PART 5: CONCLUSIONS ---
\chapter*{CONCLUSIONS}
\addcontentsline{toc}{chapter}{CONCLUSIONS}
\label{ch:conclusions}
\begingroup
\setcounter{secnumdepth}{0}
\input{chapters/conclusions}
\endgroup

{\singlespacing
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\refname}{REFERENCES}
\setlength{\bibitemsep}{\baselineskip}
\bibliographystyle{apacite_et_al}
\bibliography{references}
}

\combinedlistoftablesandfigures

\begingroup
\captionsetup{list=false}
\setcounter{table}{0}
\setcounter{figure}{0}
\setcounter{equation}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\theequation}{A\arabic{equation}}
\input{chapters/appendix/appendix_A_synthetic_control}
\endgroup

\end{document}
