% Chapter 5: Interpretation
This chapter discusses what the results in Chapter~\ref{ch:results} mean for the thesis argument.

\section{Interpreting the Main Causal Result}
\label{sec:interpretation-main-causal-result}

The result does not support the broad claim that urgent pump messages always raise returns. The clearest positive estimate for \textit{Urgency Index} appears in the small-cap stratum in Table~\ref{tab:stratified-ate-results}. Because the final stage of the Double Machine Learning procedure regresses outcome residuals onto treatment residuals, this estimate of 0.0301 can be interpreted similarly to a standard linear coefficient. In practical terms, since the outcome is a \textit{10-Minute Logarithmic Return} and the treatment is scaled as a ratio from 0 to 1, a 10 percentage point increase in the share of uppercase characters within the signal message is associated with an expected increase of approximately 0.30\% in the short-term return, holding other signal characteristics and baseline market conditions constant. The evidence nevertheless points to a conditional small-cap urgency effect rather than to a general rule for all low-valued coins.

The micro-cap result is important because it prevents a stronger reading of H1. In Chapter~\ref{ch:data-description} it was documented that the smallest tokens in the sample tend to trade infrequently, with long gaps between transactions. Chapter~\ref{ch:results} reports no significant treatment effect in the micro-cap stratum. However, this lack of statistical significance might simply be due to the smaller number of data points available for these lowest-valued assets, making it difficult for the model to isolate a clear effect. This does not mean that micro-cap pumps never react to urgent signals. It merely indicates that the current sample cannot prove such a claim. The safer interpretation is that urgency plays a positive role in the small-cap segment, while the evidence remains inconclusive for the micro-cap market.

The comparison with the other signal features also helps clarify the result. \textit{Hype Intensity} and \textit{Instruction Density} do not follow the same positive pattern as \textit{Urgency Index}. In this sample, the immediate urgency of the reveal signal is a stronger driver of the price response than the amount of hype or instructions provided beforehand. This is consistent with the idea that pump messages may function as sharp coordination triggers rather than as a product of sustained pre-reveal pressure.

\section{\textit{Market Capitalization} as a Boundary Condition}
\label{sec:interpretation-market-cap-boundary}

The most consistent pattern in Chapter~\ref{ch:results} concerns H2. The stratified estimates and the Causal Forest pattern in Figure~\ref{fig:hte-market-cap} both show that the estimated urgency effect is concentrated in lower-valued assets and becomes weaker as \textit{Market Capitalization} rises. This is in line with the theoretical argument that coordinated buying pressure should matter more in thinner markets, where sudden demand is harder to absorb \cite{amihud2002illiquidity}.

At the same time, the market-size story is not as simple as ``smaller means easier to pump''. The positive part of the pattern shows up mainly in the lower market-cap range, while the micro-cap stratum remains uncertain and the mid- and large-cap estimates turn negative. This means \textit{Market Capitalization} does not sort assets from most to least vulnerable in a clean way. At the upper end, the mid- and large-cap estimates suggest that deeper markets do not convert the signal into the same positive price response. At the lower end, micro-cap assets trade so infrequently that the available data, only 289 events, possibly could not produce a reliable estimate. In this sample, the detectable positive effect occupies the space in between, roughly the 7 million USD to 40 million USD lower market-cap range, where markets are thin enough to move but active enough to estimate.

The negative estimates for larger assets should also be interpreted with caution. They are consistent with the idea that urgent wording loses effectiveness when the market is deeper. They do not prove, however, that urgency itself mechanically lowers returns in larger assets. The DML estimator adjusts for observed pre-event conditions, but cannot account for unobserved ones, such as whether organizers systematically target different coin types or apply different signal styles in larger market segments. The result is therefore best read as a descriptive pattern. Urgency effects differ across the market-cap spectrum. It does not imply that market size structurally determines how urgency operates.

\section{Prediction and Causal Interpretation}
\label{sec:interpretation-prediction-causality}

The predictive results offer a distinct perspective, answering a fundamentally different question than the causal DML estimates. The grouped holdout XGBoost model in Table~\ref{tab:classification-no-channel} demonstrates that pre-event market variables and signal characteristics contain valuable information for determining whether a pump will produce a positive \textit{10-Minute Logarithmic Return}. This supports H3 as a forecasting claim, because the model performs effectively even when omitting \textit{Channel ID} and evaluating on a test set containing unseen channels.

These predictive features should not be read as causes by themselves. Feature importance in Figure~\ref{fig:feature-importance-no-channel} identifies variables that help the model separate successful from unsuccessful events. It does not show which variables independently caused the return. For example, \textit{Coin Age}, weekend timing, illiquidity measures, and signal features may all help prediction because they describe the kind of asset or event selected by organizers. Figure~\ref{fig:channel-returns-distribution} adds context to this point. Channels differ in their typical \textit{10-Minute Logarithmic Return} values, even though most large channels remain close to zero. Excluding \textit{Channel ID} therefore reduces direct dependence on channel history, but observable features may still carry part of the organizer's habits.

The threshold analysis in Table~\ref{tab:threshold-optimization} highlights the practical trade-offs involved in pump prediction. Raising the decision threshold improves precision, which effectively reduces a trader's exposure to ``bad guesses'' by filtering out false signals. However, this stricter criterion also lowers recall, thereby reducing the total number of available trading opportunities. This trade-off is central to the problem because pump classification is not a neutral task. A model sensitive enough to catch a larger share of successful events will inevitably produce more false positives. Consequently, while the thesis shows that pump \textit{Success} is predictable to a meaningful degree, it does not suggest that a high prediction probability guarantees a profitable outcome.

The predictive and causal parts are still connected. Taken together, they suggest a common pattern. Pump outcomes depend on the interaction between message design and earlier market conditions. The causal results isolate urgency under stronger assumptions, while the predictive model combines many signals to classify outcomes.

\section{Interpretive Limits}
\label{sec:interpretation-limits}

The first limit is the observational nature of the study. The DML framework improves adjustment for observed pre-reveal conditions, but the signals were not randomly assigned. Organizers may choose stronger wording for coins that they already expect to move, or for events where they believe participants need stronger pressure. The reported E-value for the small-cap urgency estimate in Table~\ref{tab:evalue-results} gives an approximate sensitivity check for unmeasured confounding, but it does not remove this concern.

The reliance on a 10-minute outcome window further limits the interpretation of the predictive findings. By measuring the \textit{10-Minute Logarithmic Return} shortly after the reveal, the analysis captures the immediate market reaction, but it establishes a clear distinction between model ``success'' and actual trader ``profit''. A positive \textit{10-Minute Logarithmic Return} indicates that the model has correctly identified an upward price movement. However, it does not account for the practical costs of execution, such as transaction fees, slippage, and the extreme speed required to enter and exit a position, which ultimately determine whether a participant can realize a gain. The intended takeaway is that while the thesis demonstrates that these market movements are systematically predictable, it does not imply a guarantee of individual profitability or a final picture of the event's financial outcome.

Finally, the no-channel predictive design reduces direct reliance on channel identity, but it cannot fully remove organizer patterns from the data. The secondary channel-including specification reported in Chapter~\ref{ch:results} does not improve the main predictive result, but this should not be read as proof that organizer behavior is absent. The channel-level return distribution in Figure~\ref{fig:channel-returns-distribution} shows that outcomes are not identical across Telegram groups. Channels may repeatedly choose similar coins, timings, exchanges, or message styles, and these habits can shape the information available to the model. This does not invalidate the predictive result, but it narrows its meaning. The model shows that \textit{Success} can be predicted from available event information. It does not prove that the same relationships would hold for all pump groups or settings outside the observed sample.
